Computation Issues

### 2019-08-24 journals/03-applying_lime_logistic/03-applying_lime_logistic.Rmd
worked on organizing journals

---
title: "07-logistic_regression"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

I have been thinking a lot about it is difficult to determine whether LIME is providing reasonable explanations, because we cannot know what the random forest is doing. That is the point of LIME. Heike suggested that we fit a logistic regression model to the training bullet data. We can then apply LIME to the logistic regression and determine if the explanations match what we would expect based on the coefficients of the logistic regression that we can interpret. She also suggested including interactions in the logistic regression model. This journal goes through the process of fitting several logisitic regression models, applying LIME to them, and comparing the results.

```{r}
# Load libraries
library(furrr)
library(future)
library(glmnet)
library(gretchenalbrecht)
library(hierNet)
library(lime)
library(tidyverse)

>>>>

---
title: "Applying Lime to a Logisitic Regression"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

I have been thinking a lot about it is difficult to determine whether LIME is xproviding reasonable explanations, because we cannot know what the random forest is doing. That is the point of LIME. Heike suggested that we fit a logistic regression model to the training bullet data. We can then apply LIME to the logistic regression and determine if the explanations match what we would expect based on the coefficients of the logistic regression that we can interpret. She also suggested including interactions in the logistic regression model. This journal goes through the process of fitting several logisitic regression models, applying LIME to them, and comparing the results.

```{r}
# Load libraries
library(furrr)
library(future)
library(glmnet)
library(gretchenalbrecht)
library(lime)
library(tidyverse)

### 2019-08-24 journals/03-applying_lime_rtrees/03-applying_lime_rtrees.Rmd
worked on organizing journals

---
title: "05-applying_lime_to_rtrees"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv")
```

# Description of `rtrees`

The random forest model that was fit to the Hamby training dataset is discussed in the paper "Automatic Matching of Bullet Land Impressions" by E. Hare Et. Al. The model is available through the GitHub version of the `bulletr` package. The link to the GitHub repository for `bulletr` is https://github.com/CSAFE-ISU/bulletr.

The random forest model is stored in the package `bulletr` as a data object called `rtrees`. The paper does not include many details about the model fitting process, so I looked at some of the model features, which are printed below. These values told me that the model was fit using the predictive features of `ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks` with 300 trees and the parameter of `mtry` set to 3, which is the "number of variables randomly sampled as candidates at each split" as explained in the `randomforest` package documentation.

```{r}
# Importance values from the random forest
rtrees$importance

# The number of trees used in the random forest
rtrees$ntree

# "Number of variables randomly sampled as candidates at each split"
rtrees$mtry
```

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../../data/hamby224_lime_inputs.rds")) {

@@

  # Save the lime objects
  saveRDS(hamby224_lime_inputs, "../../data/hamby224_lime_inputs.rds")
  saveRDS(hamby224_lime, "../../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../../data/hamby224_explain.rds")

@@

  # Load in the lime objects
  hamby224_lime_inputs <- readRDS("../../data/hamby224_lime_inputs.rds")
  hamby224_lime <- readRDS("../../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../../data/hamby224_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../../data/hamby224_bins.csv")) {

@@

  # Save the bin boundaries and bins
  saveRDS(hamby224_bin_boundaries, "../../data/hamby224_bin_boundaries.rds")
  saveRDS(hamby224_bins, "../../data/hamby224_bins.rds")

@@

  # Load in the bin boundaries and bins
  hamby224_bin_boundaries <- readRDS("../../data/hamby224_bin_boundaries.rds")
  hamby224_bins <- readRDS("../../data/hamby224_bins.rds")

@@

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../../data/hamby224_test_explain.rds")) {

@@

  # Save the combined test and explain data
  saveRDS(hamby224_test_explain, "../../data/hamby224_test_explain.rds")

@@

  # Load in the data
  hamby224_test_explain <- readRDS("../../data/hamby224_test_explain.rds")

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../../data/hamby224_lime_comparisons.rds")) {

@@

  # Save the lime comparison data frame
  saveRDS(hamby224_lime_comparisons, "../../data/hamby224_lime_comparisons.rds")

@@

  # Load in the lime comparison data frame
  hamby224_lime_comparisons <- readRDS("../../data/hamby224_lime_comparisons.rds")

@@

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../../data/hamby224_sensitivity_inputs.rds")) {

@@

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../../data/hamby224_sensitivity_outputs.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../../data/hamby224_sensitivity_outputs.rds")

@@

```{r}
# Join the sensativity outputs with the test data if not already saved
if(!file.exists("../../data/hamby224_sensitivity_joined.rds")) {

@@

  # Save the joined test and explained dataframes
  saveRDS(hamby224_sensitivity_joined, "../../data/hamby224_sensitivity_joined.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_joined <- readRDS("../../data/hamby224_sensitivity_joined.rds")

>>>>

---
title: "Applying lime to `rtrees`"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../../data/hamby224_test.csv")
```

@@

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("../../../data/hamby224_lime_inputs.rds")) {

@@

  # Save the lime objects
  saveRDS(hamby224_lime_inputs, "../../../data/hamby224_lime_inputs.rds")
  saveRDS(hamby224_lime, "../../../data/hamby224_lime.rds")
  saveRDS(hamby224_explain, "../../../data/hamby224_explain.rds")

@@

  # Load in the lime objects
  hamby224_lime_inputs <- readRDS("../../../data/hamby224_lime_inputs.rds")
  hamby224_lime <- readRDS("../../../data/hamby224_lime.rds")
  hamby224_explain <- readRDS("../../../data/hamby224_explain.rds")

@@

```{r}
# Create a dataframe with the bins
if (!file.exists("../../../data/hamby224_bins.csv")) {

@@

  # Save the bin boundaries and bins
  saveRDS(hamby224_bin_boundaries, "../../../data/hamby224_bin_boundaries.rds")
  saveRDS(hamby224_bins, "../../../data/hamby224_bins.rds")

@@

  # Load in the bin boundaries and bins
  hamby224_bin_boundaries <- readRDS("../../../data/hamby224_bin_boundaries.rds")
  hamby224_bins <- readRDS("../../../data/hamby224_bins.rds")

@@

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../../../data/hamby224_test_explain.rds")) {

@@

  # Save the combined test and explain data
  saveRDS(hamby224_test_explain, "../../../data/hamby224_test_explain.rds")

@@

  # Load in the data
  hamby224_test_explain <- readRDS("../../../data/hamby224_test_explain.rds")

@@

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("../../../data/hamby224_lime_comparisons.rds")) {

@@

  # Save the lime comparison data frame
  saveRDS(hamby224_lime_comparisons, "../../../data/hamby224_lime_comparisons.rds")

@@

  # Load in the lime comparison data frame
  hamby224_lime_comparisons <- readRDS("../../../data/hamby224_lime_comparisons.rds")

@@

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../../../data/hamby224_sensitivity_inputs.rds")) {

@@

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../../../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../../../data/hamby224_sensitivity_outputs.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../../../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../../../data/hamby224_sensitivity_outputs.rds")

@@

```{r}
# Join the sensativity outputs with the test data if not already saved
if(!file.exists("../../../data/hamby224_sensitivity_joined.rds")) {

@@

  # Save the joined test and explained dataframes
  saveRDS(hamby224_sensitivity_joined, "../../../data/hamby224_sensitivity_joined.rds")

@@

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_joined <- readRDS("../../../data/hamby224_sensitivity_joined.rds")

### 2019-08-24 journals/04-assessing_lime/04-assessing_lime.Rmd
worked on organizing journals

---
title: "06-assessing_lime_explanations"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

@@

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv") # created in the paper

# Read in the test_explain data
hamby224_test_explain <- readRDS("../../data/hamby224_test_explain.rds")

hamby224_sensitivity_joined <- readRDS("../../data/hamby224_sensitivity_joined.rds")

@@

```{r}
# Read in the lime comparison data (the comparisons are run in the firearm examiner
# paper since they will be discussed in the paper)
hamby224_lime_comparisons <- readRDS("../../data/hamby224_lime_comparisons.rds")
```

@@

>>>>

---
title: "Assessing Lime"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

@@

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../../data/hamby224_test.csv") # created in the paper

# Read in the test_explain data
hamby224_test_explain <- readRDS("../../../data/hamby224_test_explain.rds")

hamby224_sensitivity_joined <- readRDS("../../../data/hamby224_sensitivity_joined.rds")

@@

```{r}
# Read in the lime comparison data (the comparisons are run in the firearm examiner
# paper since they will be discussed in the paper)
hamby224_lime_comparisons <- readRDS("../../../data/hamby224_lime_comparisons.rds")
```

@@

# My LIME

### Random Forest Model

```{r}
# Fit a random forest model with the top two most important features in the model
if(!file.exists("./data/rfmodel.rds")) {

  # Fit the random forest  
  rfmodel <- randomForest(factor(samesource) ~ rough_cor + ccf, data = hamby173and252_train)

  # Save the random forest
  saveRDS(rfmodel, "./data/rfmodel.rds")

} else {

  # Load the rf model
  rfmodel <- readRDS("./data/rfmodel.rds")
}
```

```{r}
# Create a dataset with the test data input to the random forest model 
# and the predictions from the random forest model
hamby224_test_subset <- hamby224_test %>%
  na.omit() %>%
  select(case:land2, rough_cor, ccf, samesource) %>%
  mutate(rfpred = predict(rfmodel, 
                           newdata = hamby224_test %>% select(rough_cor, ccf) %>% na.omit()),
         rfscore = predict(rfmodel, 
                           newdata = hamby224_test %>% select(rough_cor, ccf) %>% na.omit(),
                           type = "prob")[,2])
```

### Applying LIME

The functions `lime` and `explain` from the lime package are applied below to the random forest model.

```{r eval = FALSE}
# Create or load the lime and explain objects for the main effects logisitic regression model
if(!file.exists("./data/lime_explain_sub.rds")) {

  # Set a seed
  set.seed(20190315)

  # Apply lime
  lime_sub <- lime(x = hamby173and252_train %>% select(rough_cor, ccf), 
                   model = as_classifier(rfmodel))

  # Apply explain
  explain_sub <- explain(x = hamby224_test %>%
                             select(rough_cor, ccf) %>% 
                             na.omit(), 
                           explainer = lime_sub, 
                           n_labels = 1,
                           n_features = 2)

  # Join the lime and explain objects in a list
  lime_explain_sub <- list(lime = lime_sub, explain = explain_sub)

  # Save the lime and explain objects
  saveRDS(lime_explain_sub, "./data/lime_explain_sub.rds")

} else {

  # Load the lime and explain objects
  lime_explain_sub <- readRDS("./data/lime_explain_sub.rds")

}
```

```{r eval = FALSE}
create_bin_df <- function(bin_cut) {
 data.frame(quantile = c("0%", "25%", "50%", "75%", "100%"), 
            bin_cut) 
}

bin_cuts <- purrr::map_df(lime_explain_sub$lime$bin_cuts, 
                          create_bin_df,
                          .id = "feature") %>%
  spread(key = feature, value = bin_cut)

bin_cuts
```

### Visualizations of Data, Model, and LIME

```{r eval = FALSE}
(case1 <- lime_explain_sub$explain %>% filter(case == 1))
```

```{r eval = FALSE}
ggplot() +
  geom_point(data = hamby173and252_train %>% filter(samesource == FALSE),
             aes(x = rough_cor, y = ccf),
             color = "darkgrey",
             alpha = 0.2) +
  geom_point(data = hamby173and252_train %>% filter(samesource == TRUE),
             aes(x = rough_cor, y = ccf),
             color = "darkorange") + 
  geom_vline(xintercept = bin_cuts$rough_cor) + 
  geom_hline(yintercept = bin_cuts$ccf) +
  theme_minimal()
```

```{r eval = FALSE}
ggplot() + 
  geom_point(data = hamby224_test_subset, 
             aes(x = rough_cor, y = ccf, color = rfscore),
             alpha = 0.5) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  geom_point(data = lime_explain_sub$explain %>% 
               filter(case == 1) %>%
               select(feature, feature_value) %>%
               spread(key = feature, value = feature_value),
             aes(x = rough_cor, y = ccf)) +
  geom_vline(xintercept = bin_cuts$rough_cor) + 
  geom_hline(yintercept = bin_cuts$ccf) + 
  theme_minimal()
```

### 2019-08-24 journals/05-iris_comparison/05-iris_comparison.Rmd
worked on organizing journals

---
title: "08-iris_comparison"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

>>>>

---
title: "Iris Comparison"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

### 2019-08-24 journals/06-other_explainers/06-other_explainers.Rmd
worked on organizing journals

---
title: "10-other_explainers"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

>>>>

---
title: "Other Explainers"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

### 2019-08-24 journals/07-computation_issues/07-computation_issues.Rmd
worked on organizing journals

---
title: "03-computation_issues"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

>>>>

---
title: "Computation Issues"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
